---
title: "2025_Regression_and_Classification"
author: "Robert Feldstein, Melissa Eckert"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# Necessary Libraries
library(readxl)
library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)
library(pROC)
```

## Data Cleaning & Importation

```{r}
# We need both the 2025 data, and the procedure data

df25 <- read_xlsx("../Data/2025Data.xlsx", sheet = 1)
dfproc <- read_xlsx("../Data/FinalProcedures.xlsx", sheet = 1)

# We need to aggregate the procedure data to the student level
# For now, we will just look at Ovarioectomy/Ovariohysterectomy and Castration

neuters <- dfproc %>%
  filter(skill %in% c("Ovarioectomy/Ovariohysterectomy", "Castration")) %>%
  group_by(ID, skill) %>%
  summarise(total = sum(total_number_performed, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = skill, values_from = total, values_fill = 0, 
              names_prefix = "Total_")

# Rename neuters column names
colnames(neuters) <- c("ID", "Total_Ov", "Total_Castration")

# Make a skill_category dataset

skill_cat <- dfproc %>%
  group_by(ID, skill_category) %>%
  summarise(total = sum(total_number_performed, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = skill_category, values_from = total, values_fill = 0, 
              names_prefix = "Total_")

# Now we need to merge the two dataframes
# Make a new column in df25 called TotalOv and TotalCa
# It looks at the ID and then pulls the total from the neuters dataframe

df25 <- df25 %>%
  left_join(neuters, by = c("UniqueID" = "ID")) %>%
  mutate(
    Total_Castration = replace_na(Total_Castration, 0),
    Total_Ov = replace_na(Total_Ov, 0)
  )

# Now we need to merge the skill_cat dataframe

df25 <- df25 %>%
  left_join(skill_cat, by = c("UniqueID" = "ID")) %>%
  mutate(
    `Total_Surgical Skills` = replace_na(`Total_Surgical Skills`, 0),
    `Total_Procedural Skills` = replace_na(`Total_Procedural Skills`, 0),
    `Total_Dental Skills` = replace_na(`Total_Dental Skills`, 0),
    `Total_Anesthesia Skills` = replace_na(`Total_Anesthesia Skills`, 0)
  )

# Rename columns to add underscores
df25 <- df25 %>%
  rename_with(~str_replace_all(., " ", "_"))

# Add time variable

labels <- c("A1", "A2", "A3", "A4", "A5", "A6", "A7", "B1", "B2", "B3", "B4", 
            "B5", "B6", "C1", "C2", "C3", "C4", "C5", "C6", "D1","D2", 
            "D3", "D4", "D5", "D6")

# Create a mapping of labels to integers
label_mapping <- setNames(1:length(labels), labels)

# Convert strBlock column to integers
df25$time<- label_mapping[df25$strBlock]

```


```{r}

# This maps missing catalog numbers to the last service for each student

# 1. Build a lookup table for the last rotation block for each student in df25
last_rotation <- df25 %>%
  group_by(UniqueID) %>%
  summarise(last_service = last(service), .groups = "drop")

# 2. For procedure data, join the last rotation block and use it for missing catalog_number
dfproc_course <- dfproc %>%
  filter(skill %in% c("Ovarioectomy/Ovariohysterectomy", "Castration")) %>%
  # join last rotation info based on student ID (which is called ID in dfproc)
  left_join(last_rotation, by = c("ID" = "UniqueID")) %>%
  # if catalog_number is missing, use the last_service from df25
  mutate(catalog_number_fixed = if_else(is.na(catalog_number), last_service, catalog_number)) %>%
  # select the updated column name
  select(ID, catalog_number = catalog_number_fixed, skill, total_number_performed)

# 3. Aggregate by student (ID) and course (catalog_number)
neuters <- dfproc_course %>%
  group_by(ID, catalog_number, skill) %>%
  summarise(total = sum(total_number_performed, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = skill, values_from = total, values_fill = 0, 
              names_prefix = "Total_")

colnames(neuters) <- c("ID", "catalog_number", "Ov", "Castration")

# 4. Merge with df25 using both UniqueID and service (which now matches catalog_number)
df25 <- df25 %>%
  left_join(neuters, by = c("UniqueID" = "ID", "service" = "catalog_number")) %>%
  mutate(
    Castration = replace_na(Castration, 0),
    Ov = replace_na(Ov, 0)
  )

# 5. Accumulate procedure totals up to the rotation block for each student
df25$CastrationPerformed <- ave(df25$Castration, df25$UniqueID, FUN = cumsum)
df25$OvPerformed <- ave(df25$Ov, df25$UniqueID, FUN = cumsum)

# If you want to apply a similar process for skill categories, do the same:
skill_cat <- dfproc %>%
  left_join(last_rotation, by = c("ID" = "UniqueID")) %>%
  mutate(catalog_number_fixed = if_else(is.na(catalog_number), last_service, catalog_number)) %>%
  select(ID, catalog_number = catalog_number_fixed, skill_category, total_number_performed) %>%
  group_by(ID, catalog_number, skill_category) %>%
  summarise(total = sum(total_number_performed, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = skill_category, values_from = total, values_fill = 0)

df25 <- df25 %>% 
  left_join(skill_cat, by = c("UniqueID" = "ID", "service" = "catalog_number")) %>%
  mutate(
    `Surgical Skills` = replace_na(`Surgical Skills`, 0),
    `Procedural Skills` = replace_na(`Procedural Skills`, 0),
    `Dental Skills` = replace_na(`Dental Skills`, 0),
    `Anesthesia Skills` = replace_na(`Anesthesia Skills`, 0)
  )

df25$Surgical_Per <- ave(df25$`Surgical Skills`, df25$UniqueID, FUN = cumsum)
df25$Procedural_Per <- ave(df25$`Procedural Skills`, df25$UniqueID, FUN = cumsum)
df25$Dental_Per <- ave(df25$`Dental Skills`, df25$UniqueID, FUN = cumsum)
df25$Anesthesia_Per <- ave(df25$`Anesthesia Skills`, df25$UniqueID, FUN = cumsum)
```


## Regression Models

```{r}

# Start with linear regression 

lm.medical <- lm(medical ~ PIPOverall + time + CastrationPerformed + 
               OvPerformed + Surgical_Per + Procedural_Per + Dental_Per + Anesthesia_Per, data = df25)
summary(lm.medical)


```

```{r}

lm.clinical <- lm(clinical_reasoning ~ PIPOverall + time + CastrationPerformed + 
               OvPerformed + Surgical_Per + Procedural_Per + Dental_Per + Anesthesia_Per, data = df25)

summary(lm.clinical)


```

Unfortunately, it does not appear to be the case that the new procedural data is significantly associated with scores. This could be due to the fact that we are working with an incomplete dataset. 

```{r}
lm.prof <- lm(professionalism ~ PIPOverall + time + CastrationPerformed + 
               OvPerformed + Surgical_Per + Procedural_Per + Dental_Per + Anesthesia_Per, data = df25)
summary(lm.prof)
```

```{r}

lm.coll <- lm(collaboration ~ PIPOverall + time + CastrationPerformed + 
               OvPerformed + Surgical_Per + Procedural_Per + Dental_Per + Anesthesia_Per, data = df25)

summary(lm.coll)

```


```{r}
# We can also do the standard time based linear regressions we did previously

time_regression <- function(df, outcome) {
  # Build the formula from the outcome string
  fmla <- as.formula(paste(outcome, "~ time"))
  fit <- lm(fmla, data = df)
  sfit <- summary(fit)
  
  # Extract the intercept and its p-value
  int_est <- coef(sfit)["(Intercept)", "Estimate"]
  int_p   <- coef(sfit)["(Intercept)", "Pr(>|t|)"]
  
  # Extract the coefficient for time and its p-value
  time_est <- coef(sfit)["time", "Estimate"]
  time_p   <- coef(sfit)["time", "Pr(>|t|)"]
  
  return(c(Intercept = int_est, Intercept_p = int_p,
           Time_coef = time_est, Time_p = time_p))
}


outcomes <- c("medical", "clinical_reasoning", "professionalism", "collaboration",
              "patient_care", "procedural", "oral", "written", "animal_population",
              "evidence", "financial", "public_health", "service_specific")

# Build a table by applying time_regression to each outcome
result_table <- do.call(rbind, lapply(outcomes, function(o) time_regression(df25, o)))
result_table <- data.frame(Outcome = outcomes, result_table)
result_table


```


```{r}
# Run it on procedural 

# lm.procedural <- lm(procedural ~  CastrationPerformed + 
              # OvPerformed + time, data = df25)

#summary(lm.procedural)

# Try it again with just the other variables 
# 
lm.procedural <- lm(procedural ~ time+ Surgical_Per + Procedural_Per + Dental_Per + Anesthesia_Per, data = df25)
summary(lm.procedural)
```





## Classification Models

First, let's repeat the 2024 logistic regression on the 2025 data 
```{r}
#Convert attribute scores into competency categories
df25_competency<- df25 |>
    mutate(across(outcomes, ~ factor(
      ifelse(. %in% c(1,2,3), 0, 1), levels = c(0,1)) )) |>
  select(strBlock, time, everything())

#function similar to above for logistic regression
time_logistic_regression <- function(df, outcome) {
  # Build the formula from the outcome string
  fmla <- as.formula(paste(outcome, "~ time"))
  fit <- glm(fmla, data = df, family = "binomial")
  sfit <- summary(fit)
  
  # Extract the intercept and its p-value
  int_est <- coef(sfit)["(Intercept)", "Estimate"]
  int_p   <- coef(sfit)["(Intercept)", "Pr(>|z|)"]
  
  # Extract the coefficient for time and its p-value
  time_est <- coef(sfit)["time", "Estimate"]
  time_p   <- coef(sfit)["time", "Pr(>|z|)"]
  
  # Obtain odds ratio for the time coefficient 
  time_odds_ratio<- exp(time_est)
  
  #Obtain model AUC
  pred<- predict(fit, type = "response")
  roc<- roc(df[[outcome]], pred)
  auc<- auc(roc)
  
  return(c(Intercept = int_est, Intercept_p = int_p,
           Time_coef = time_est, Time_odds_ratio = time_odds_ratio, Time_p = time_p, model_AUC = auc))
}

# Build a table by applying function to each outcome
result_table_log <- do.call(rbind, lapply(outcomes, function(o) time_logistic_regression(df25_competency, o)))
result_table_log <- data.frame(Outcome = outcomes, result_table_log)
result_table_log



```
Now, lets incorporate procedural variables

```{r}

#Look at medical knowledge only 

glm.medical <- glm(medical ~ PIPOverall + time + CastrationPerformed + 
               OvPerformed + Surgical_Per + Procedural_Per + Dental_Per + Anesthesia_Per, family = "binomial",data = df25_competency)
summary(glm.medical)

#Look at clinical reasoning only 
glm.clinical <- glm(clinical_reasoning ~ PIPOverall + time + CastrationPerformed + 
               OvPerformed + Surgical_Per + Procedural_Per + Dental_Per + Anesthesia_Per, family = "binomial",data = df25_competency)
summary(glm.clinical)

#Create another function to include procedural variables for all attributes  

procedure_logistic_regression <- function(df, outcome) {
  # Build the formula from the outcome string
  fmla <- as.formula(paste(outcome, "~ PIPOverall + time + CastrationPerformed + OvPerformed + Surgical_Per + Procedural_Per + Dental_Per + Anesthesia_Per"))
  fit <- glm(fmla, data = df, family = "binomial")
  sfit <- summary(fit)
  
  # Extract each coefficient and its p-value
  int_est <- coef(sfit)["(Intercept)", "Estimate"]
  int_p   <- coef(sfit)["(Intercept)", "Pr(>|z|)"]
  
  # Extract the coefficient for time and its p-value
  time_est <- coef(sfit)["time", "Estimate"]
  time_p   <- coef(sfit)["time", "Pr(>|z|)"]
  
  # Extract the coefficient for PIP and its p-value
  pip_est <- coef(sfit)["PIPOverall", "Estimate"]
  pip_p   <- coef(sfit)["PIPOverall", "Pr(>|z|)"]
  
  
  
  # Extract the coefficient for castrations and its p-value
  castration_est <- coef(sfit)["CastrationPerformed", "Estimate"]
  castration_p   <- coef(sfit)["CastrationPerformed", "Pr(>|z|)"]
  
  # Extract the coefficient for ovariectomies and its p-value
  ov_est <- coef(sfit)["OvPerformed", "Estimate"]
  ov_p   <- coef(sfit)["OvPerformed", "Pr(>|z|)"]
  
  # Extract the coefficient for surgical skills and its p-value
  surgical_est <- coef(sfit)["Surgical_Per", "Estimate"]
  surgical_p   <- coef(sfit)["Surgical_Per", "Pr(>|z|)"]
  
  # Extract the coefficient for procedural and its p-value
  procedural_est <- coef(sfit)["Procedural_Per", "Estimate"]
  procedural_p   <- coef(sfit)["Procedural_Per", "Pr(>|z|)"]
  
  # Extract the coefficient for dental and its p-value
  dental_est <- coef(sfit)["Dental_Per", "Estimate"]
  dental_p   <- coef(sfit)["Dental_Per", "Pr(>|z|)"]
  
  # Extract the coefficient for anesthesia and its p-value
  anesthesia_est <- coef(sfit)["Anesthesia_Per", "Estimate"]
  anesthesia_p   <- coef(sfit)["Anesthesia_Per", "Pr(>|z|)"]
  
  
  # Obtain odds ratio for the time coefficient 
  time_odds<- exp(time_est)
  pip_odds<- exp(pip_est)
  castration_odds<- exp(castration_est)
  ov_odds<- exp(ov_est)
  surgical_odds<-exp(surgical_est)
  procedural_odds<- exp(procedural_est)
  dental_odds<- exp(dental_est)
  anesthesia_odds<-exp(anesthesia_est)
  
  #Obtain model AUC
  pred<- predict(fit, type = "response")
  roc<- roc(df[[outcome]], pred)
  auc<- auc(roc)
  
  return(c(Intercept = int_est, Intercept_p = int_p, PIP = pip_est, PIP_odds_ratio = pip_odds, PIP_p = pip_p,
           Time_coef = time_est, Time_odds_ratio = time_odds, Time_p = time_p,
           Castration_coef = castration_est, Castration_odds_ratio = castration_odds, Castration_p = castration_p,
           Ov_coef = ov_est, Ov_odds_ratio = ov_odds, Ov_p  = ov_p, 
           Surgical_coef = surgical_est, Surgical_odds_ratio = surgical_odds, Surgical_p = surgical_p,
           Procedural_coef = procedural_est, Procedural_odds_ratio = procedural_odds, Procedural_p = procedural_p,
           Dental_coef = dental_est, Dental_odds_ratio = dental_odds, Dental_p = dental_p,
           Anesthesia_coef = anesthesia_est, Anesthesia_odds_ratio = anesthesia_odds, Anesthesia_p = anesthesia_p,model_AUC = auc))
}

# Build a table by applying function to each outcome
result_table_procedural <- do.call(rbind, lapply(outcomes, function(o) procedure_logistic_regression(df25_competency, o)))
result_table_procedural <- data.frame(Outcome = outcomes, result_table_procedural)
result_table_procedural

```

