---
title: "Percentage Change Regressions"
author: "Robert Feldstein, Melissa Eckert"
date: "2025-02-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Linear Regression

We will be repeating a set of linear regressions performed by the client, 
but with the expressed interest of examining the percentage change in scores.

```{r}

library(readxl)
library(dplyr)
df <- read_excel("Data/FinalDataSet.xlsx")

labels <- c("A1", "A2", "A3", "A4", "A5", "A6", "A7", "B1", "B2", "B3", "B4", 
            "B5", "B6", "C1", "C2", "C3", "C4", "C5", "C6", "D1","D2", 
            "D3", "D4", "D5", "D6")

# Create a mapping of labels to integers
label_mapping <- setNames(1:length(labels), labels)

# Convert strBlock column to integers
df$strBlockint<- label_mapping[df$strBlock]


```

```{r}

# Make a simpler DF to look at the four main scoring categories

base_df <- df[,c("UniqueID", "strBlock", "strBlockint", "medical", "clinical_reasoning", 
                 "professionalism", "collaboration", "Hosp",
                 "Early", "PIPYN", "PIPActive")]

```



```{r}
# Original Linear regressions
lm.medical <- lm(medical ~ strBlockint, data = base_df)
summary(lm.medical)

lm.clinical <- lm(clinical_reasoning ~ strBlockint, data = base_df)
summary(lm.clinical)

lm.professionalism <- lm(professionalism ~ strBlockint, data = base_df)
summary(lm.professionalism)

lm.collaboration <- lm(collaboration ~ strBlockint, data = base_df)
summary(lm.collaboration)

lm.patient_care <- lm(patient_care ~ strBlockint, data = df)
summary(lm.patient_care)

lm.procedural_skills <- lm(procedural ~ strBlockint, data = df)
summary(lm.procedural_skills)

lm.oral_comm <- lm(oral~ strBlockint, data = df)
summary(lm.oral_comm)

lm.written_comm <- lm(written ~ strBlockint, data = df)
summary(lm.written_comm)


lm.pop_care <- lm(animal_population ~ strBlockint, data = df)
summary(lm.pop_care)

lm.evidence <- lm(evidence ~ strBlockint, data = df)
summary(lm.evidence)

lm.financial <- lm(financial ~ strBlockint, data = df)
summary(lm.financial)

lm.public <- lm(public_health ~ strBlockint, data = df)
summary(lm.public)

lm.service <- lm(service_specific ~ strBlockint, data = df)
summary(lm.service)


```

```{r}

# Look at log changes in scores to be more easily comparable

log.overall <- lm(log(Average) ~ strBlockint, data = df)
summary(log.overall)

log.medical <- lm(log(medical) ~ strBlockint, data = base_df)
summary(log.medical)

log.clinical <- lm(log(clinical_reasoning) ~ strBlockint, data = base_df)
summary(log.clinical)

log.professionalism <- lm(log(professionalism) ~ strBlockint, data = base_df)
summary(log.professionalism)

log.collaboration <- lm(log(collaboration) ~ strBlockint, data = base_df)
summary(log.collaboration)

log.patient_care <- lm(log(patient_care) ~ strBlockint, data = df)
summary(log.patient_care)

log.procedural_skills <- lm(log(procedural) ~ strBlockint, data = df)
summary(log.procedural_skills)

log.oral_comm <- lm(log(oral) ~ strBlockint, data = df)
summary(log.oral_comm)

log.written_comm <- lm(log(written) ~ strBlockint, data = df)
summary(log.written_comm)

log.pop_care <- lm(log(animal_population) ~ strBlockint, data = df)
summary(log.pop_care)

log.evidence <- lm(log(evidence) ~ strBlockint, data = df)
summary(log.evidence)

log.financial <- lm(log(financial) ~ strBlockint, data = df)
summary(log.financial)

log.public <- lm(log(public_health) ~ strBlockint, data = df)
summary(log.public)

log.service <- lm(log(service_specific) ~ strBlockint, data = df)
summary(log.service)



```

```{r}
# Collect all the coefficients and take their exponent to get the percentage change

overall_pct <- (exp(coef(log.overall))[2]-1)*100
med_pct <- (exp(coef(log.medical))[2]-1)*100
cli_pct <- (exp(coef(log.clinical))[2]-1)*100
pro_pct <- (exp(coef(log.professionalism))[2]-1)*100
col_pct <- (exp(coef(log.collaboration))[2]-1)*100
pat_pct <- (exp(coef(log.patient_care))[2]-1)*100
proc_pct <- (exp(coef(log.procedural_skills))[2]-1)*100
oral_pct <- (exp(coef(log.oral_comm))[2]-1)*100
writ_pct <- (exp(coef(log.written_comm))[2]-1)*100
pop_pct <- (exp(coef(log.pop_care))[2]-1)*100
evid_pct <- (exp(coef(log.evidence))[2]-1)*100
fin_pct <- (exp(coef(log.financial))[2]-1)*100
pub_pct <- (exp(coef(log.public))[2]-1)*100
serv_pct <- (exp(coef(log.service))[2]-1)*100

pcts <- c(overall_pct, med_pct, cli_pct, pro_pct, col_pct, pat_pct, proc_pct, oral_pct, writ_pct, pop_pct, evid_pct, fin_pct, pub_pct, serv_pct)

# Collect information on the intercept terms

overall_int <- exp(coef(log.overall))[1]
med_int <- exp(coef(log.medical))[1]
cli_int <- exp(coef(log.clinical))[1]
pro_int <- exp(coef(log.professionalism))[1]
col_int <- exp(coef(log.collaboration))[1]
pat_int <- exp(coef(log.patient_care))[1]
proc_int <- exp(coef(log.procedural_skills))[1]
oral_int <- exp(coef(log.oral_comm))[1]
writ_int <- exp(coef(log.written_comm))[1]
pop_int <- exp(coef(log.pop_care))[1]
evid_int <- exp(coef(log.evidence))[1]
fin_int <- exp(coef(log.financial))[1]
pub_int <- exp(coef(log.public))[1]
serv_int <- exp(coef(log.service))[1]


ints <- c(overall_int,med_int, cli_int, pro_int, col_int, pat_int, proc_int, oral_int, writ_int, pop_int, evid_int, fin_int, pub_int, serv_int)

# Calculate the expected final score for each category after 25 rotations
# final = (1+rate)^25 * initial

overall_final <- overall_int * (1 + overall_pct/100)^25
med_final <- med_int * (1 + med_pct/100)^25
cli_final <- cli_int * (1 + cli_pct/100)^25
pro_final <- pro_int * (1 + pro_pct/100)^25
col_final <- col_int * (1 + col_pct/100)^25
pat_final <- pat_int * (1 + pat_pct/100)^25
proc_final <- proc_int * (1 + proc_pct/100)^25
oral_final <- oral_int * (1 + oral_pct/100)^25
writ_final <- writ_int * (1 + writ_pct/100)^25
pop_final <- pop_int * (1 + pop_pct/100)^25
evid_final <- evid_int * (1 + evid_pct/100)^25
fin_final <- fin_int * (1 + fin_pct/100)^25
pub_final <- pub_int * (1 + pub_pct/100)^25
serv_final <- serv_int * (1 + serv_pct/100)^25

finals <- c(overall_final,med_final, cli_final, pro_final, col_final, pat_final, proc_final, oral_final, writ_final, pop_final, evid_final, fin_final, pub_final, serv_final)

# Put this information into a table

categories <- c("Overall", "Medical", "Clinical Reasoning", "Professionalism", "Collaboration", "Patient Care", "Procedural Skills", "Oral Communication", "Written Communication", "Population Health", "Evidence-based Medicine", "Financial Literacy", "Public Health", "Service Specific")

pct_change <- data.frame("Category" = categories,
                         "PercentageGrowthPerRotation" = pcts,
                         "ExpectedStartingScore" = ints,
                         "ExpectedFinalScore" = finals)                      

pct_change <- pct_change %>% 
 mutate_if(is.numeric, round,digits = 3)

pct_change

```

Here we see by taking a log-transformation the medical and clinical reasoning score categories grow almost twice as quickly as scores in professionalism and collaboration. This is slightly different from the original linear regressions, which would not emphasize the difference in growth rates between the categories.


```{r}
# Convert this table to a nicer looking table
library(kableExtra)
pct_change %>%
  kable("html") %>%
  kable_styling("striped", full_width = F) %>% 
  column_spec(1, bold = T) %>%
  column_spec(2:4, width = "10%")

```



```{r}

# Add %-change to the original table

base_df <- base_df %>%
  group_by(UniqueID) %>%  # Ensure percentage change is calculated within each student
  mutate(medical_pct = (medical - lag(medical)) / lag(medical) * 100, 
         clinical_pct = (clinical_reasoning - lag(clinical_reasoning)) / lag(clinical_reasoning) * 100,
         professionalism_pct = (professionalism - lag(professionalism)) / lag(professionalism) * 100,
         collaboration_pct = (collaboration - lag(collaboration)) / lag(collaboration) * 100)

base_df <- base_df %>%
  mutate(medical_pct = ifelse(is.na(medical_pct), 0, medical_pct),
         clinical_pct = ifelse(is.na(clinical_pct), 0, clinical_pct),
         professionalism_pct = ifelse(is.na(professionalism_pct), 0, professionalism_pct),
         collaboration_pct = ifelse(is.na(collaboration_pct), 0, collaboration_pct))

# Now do regressions on these percentage changes

lm.medical_pct <- lm(medical_pct ~ strBlockint, data = base_df)
summary(lm.medical_pct)

lm.clinical_pct <- lm(clinical_pct ~ strBlockint, data = base_df)
summary(lm.clinical_pct)

lm.professionalism_pct <- lm(professionalism_pct ~ strBlockint, data = base_df)
summary(lm.professionalism_pct)

lm.collaboration_pct <- lm(collaboration_pct ~ strBlockint, data = base_df)
summary(lm.collaboration_pct)




```

These regressions are not significant at all, and probably should be discarded.

