---
title: "Percentage Change Regressions"
author: "Robert Feldstein, Melissa Eckert"
date: "2025-02-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Linear Regression

We will be repeating a set of linear regressions performed by the client, 
but with the expressed interest of examining the percentage change in scores.

```{r}

library(readxl)
library(dplyr)
df <- read_excel("Data/FinalDataSet.xlsx")

labels <- c("A1", "A2", "A3", "A4", "A5", "A6", "A7", "B1", "B2", "B3", "B4", 
            "B5", "B6", "C1", "C2", "C3", "C4", "C5", "C6", "D1","D2", 
            "D3", "D4", "D5", "D6")

# Create a mapping of labels to integers
label_mapping <- setNames(1:length(labels), labels)

# Convert strBlock column to integers
df$strBlockint<- label_mapping[df$strBlock]


```

```{r}

# Make a simpler DF to look at the four main scoring categories

base_df <- df[,c("UniqueID", "strBlock", "strBlockint", "medical", "clinical_reasoning", 
                 "professionalism", "collaboration", "Hosp",
                 "Early", "PIPYN", "PIPActive")]

```



```{r}
# Original Linear regressions
lm.medical <- lm(medical ~ strBlockint, data = base_df)
summary(lm.medical)

lm.clinical <- lm(clinical_reasoning ~ strBlockint, data = base_df)
summary(lm.clinical)

lm.professionalism <- lm(professionalism ~ strBlockint, data = base_df)
summary(lm.professionalism)

lm.collaboration <- lm(collaboration ~ strBlockint, data = base_df)
summary(lm.collaboration)


```

```{r}

# Look at log changes in scores to be more easily comparable

log.medical <- lm(log(medical) ~ strBlockint, data = base_df)
summary(log.medical)

log.clinical <- lm(log(clinical_reasoning) ~ strBlockint, data = base_df)
summary(log.clinical)

log.professionalism <- lm(log(professionalism) ~ strBlockint, data = base_df)
summary(log.professionalism)

log.collaboration <- lm(log(collaboration) ~ strBlockint, data = base_df)
summary(log.collaboration)



```

```{r}
# Collect all the coefficients and take their exponent to get the percentage change

med_pct <- (exp(coef(log.medical))[2]-1)*100
cli_pct <- (exp(coef(log.clinical))[2]-1)*100
pro_pct <- (exp(coef(log.professionalism))[2]-1)*100
col_oct <- (exp(coef(log.collaboration))[2]-1)*100

# Put this information into a table


pct_change <- data.frame("Category" = c("Medical", "Clinical Reasoning", "Professionalism", "Collaboration"),
                         "PercentageGrowthPerRotation" = c(med_pct, cli_pct, pro_pct, col_oct))                        

pct_change
```

Here we see by taking a log-transformation the medical and clinical reasoning score categories grow almost twice as quickly as scores in professionalism and collaboration. This is slightly different from the original linear regressions, which would not emphasize the difference in growth rates between the categories.





