---
title: "Replication"
author: "Robert Feldstein, Melissa Eckert, Yue Li, Kanna Chen"
date: "2025-02-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(openxlsx)

```

Load in the dataset
```{r}
# Open the dataset
df <- read.xlsx("../Data/FinalDataSet.xlsx")
```

## Entrustment Score Distribution

```{r}

score_columns <- colnames(df)[4:16]
# Store all of the scores in a single vector
all_scores <- c(rep(NA, length(score_columns)))
for (i in 1:length(score_columns)) {
  all_scores <- c(all_scores, df[[score_columns[i]]])
}

# Remove NA values 
all_scores <- all_scores[!is.na(all_scores)]

mean_score <- mean(all_scores)

# Add the mean of the score as text to the plot
ggplot(data = data.frame(score = all_scores), aes(x = score)) +
  geom_histogram(binwidth = 1, fill = "red", color = "black") +
  labs(title = "Distribution of Entrustment Scores",
       x = "Entrustment Score",
       y = "Frequency") +
  theme_minimal() +
  geom_vline(xintercept = mean_score, color = "blue") +
  annotate("text", x = 1, y = 7000, label = paste("Mean Score: ", round(mean_score, 2))) +
  annotate("text", x = 1.5, y = 6000, label = paste("Total Observations: ", length(all_scores)))




```

```{r}
# Test linear regression
sample_df <- df[,c("strBlock", "medical", "clinical_reasoning", "professionalism", "collaboration", "patient_care")]

# Categorical Model
cat.fit <- lm(medical ~ strBlock, data = sample_df)


# Linear Model
# Convert strBlock to numeric 
length(unique(sample_df$strBlock))

labels <- c("A1", "A2", "A3", "A4", "A5", "A6", "A7", "B1", "B2", "B3", "B4", 
            "B5", "B6", "C1", "C2", "C3", "C4", "C5", "C6", "D1","D2", 
            "D3", "D4", "D5", "D6")

# Create a mapping of labels to integers
label_mapping <- setNames(1:length(labels), labels)

# Convert strBlock column to integers
sample_df$strBlockint<- label_mapping[sample_df$strBlock]


# Make medical linear regression
lm.fit.medical <- lm(medical ~ strBlockint, data = sample_df)
summary(lm.fit.medical)

# Make clinical reasoning linear regression
lm.fit.clinical_reasoning <- lm(clinical_reasoning ~ strBlockint, data = sample_df)
summary(lm.fit.clinical_reasoning)

# Make professionalism linear regression
lm.fit.professionalism <- lm(professionalism ~ strBlockint, data = sample_df)
summary(lm.fit.professionalism)

# Make collaboration linear regression
lm.fit.collaboration <- lm(collaboration ~ strBlockint, data = sample_df)
summary(lm.fit.collaboration)

# Make patient care linear regression
lm.fit.patient_care <- lm(patient_care ~ strBlockint, data = sample_df)
summary(lm.fit.patient_care)


```
```{r}
# Get the 25th percentile and 10th percentile of each of the scores

# Manipulating Time

columns <- c("medical", "clinical_reasoning", "professionalism", "collaboration", "patient_care")
percentiles <- c(0.25, 0.10)
nmat <- matrix(NA, nrow = length(columns), ncol = length(percentiles))

for (i in 1:length(columns)) {
  for (j in 1:length(percentiles)) {
    nmat[i, j] <- quantile(df[[columns[i]]], probs = percentiles[j], na.rm=T)
  }
}

rownames(nmat) <- columns
colnames(nmat) <- paste0(percentiles * 100, "%")
nmat


```




```{r}
# Make residual plots for each of the scores
# Medical residual plot
plot(lm.fit.medical, which = 1)


```


```{r}

# Comparison of Medical Knowledge Regression Curves with Respect to Professional Improvement Plans

# Group 1: Students who never joined a PIP up to block ? 

# Group 2: Students prior to joining a PIP at block?

# Group 3: Students who never joined a PIP after block ?

# Group 4: Students after joining a PIP at block ?

comparison <- df[,c("UniqueID", "strBlock", "medical", "PIPYN", "PIPActive")]
comparison$time<- label_mapping[comparison$strBlock]
grouped <- comparison %>% group_by(strBlock) %>% summarise(sum(PIPActive))

barplot(grouped$`sum(PIPActive)`, names.arg = grouped$strBlock, col = "blue", main = "PIP Active by Block", xlab = "Block", ylab = "Number of Students")


```

```{r}
# We'd really like to make a column called "PIPStart" that is 1 if the student started a PIP in that block and 0 otherwise

comparison <- comparison %>% group_by(UniqueID) %>% mutate(PIPStart = ifelse(PIPActive == 1 & lag(PIPActive) == 0, 1, 0))

# Now make a bar chart of the number of students who started a PIP in each block
grouped <- comparison %>% group_by(strBlock) %>% summarise(sum(PIPStart))

barplot(grouped$`sum(PIPStart)`, names.arg = grouped$strBlock, col = "blue", main = "PIP Start by Block", xlab = "Block", ylab = "Number of Students")


```


## Fixed Effects Model

```{r}
# What is the best way to analyze the PIP data with student scores, given that students start PIPs at different times?

library(fixest) # panel regression

model <- feols(medical ~ PIPActive | UniqueID + strBlock, data = comparison)

summary(model)



```


Students who joined a PIP are exected to have improved their medical scores by 0.088 points compared to had they not joined a PIP. This is not statistically significant at a 0.05 level. The test suggests that PIPs marginally improved student scores, but the test's results cannot be considered statistically significant. 


```{r}
ggplot(df, aes(x = factor(medical), fill = factor(PIPActive))) +
  geom_bar(position = "fill") +
  labs(x = "Score", y = "Proportion", fill = "PIP Active") +
  theme_minimal()
```


## Are PIPs Effective?

First start by assuming that most PIPs start in A7. 

```{r}

# The Research Question would be as Follows 
# Would students enrolled in a PIP have lower scores than had they not enrolled in a PIP?
library(fixest)


# Make a column called Post which is after A7

df$Post <- df$strBlockInt


did_model <- feols(Average ~ PIPYN * PIPActive | UniqueID + service + Hosp, data = df)
summary(did_model)



```

## Comparison of Entrustment Scores between Student Employees & Student Non-employees
```{r}
#simple linear regression
df2<- df |>
  select(strBlock, Hosp, Average) |>
  mutate(Hosp = factor(Hosp, labels = c("No Hospital Job", "Hospital Job")))

labels <- c("A1", "A2", "A3", "A4", "A5", "A6", "A7", "B1", "B2", "B3", "B4", 
            "B5", "B6", "C1", "C2", "C3", "C4", "C5", "C6", "D1","D2", 
            "D3", "D4", "D5", "D6")

# create a mapping of labels to integers
label_mapping <- setNames(1:length(labels), labels)

# convert strBlock column to integers
df2$strBlockint<- label_mapping[df2$strBlock]

#Without considering strBlock as a predictor
fit<- lm(Average ~ Hosp, data = df2)
summary(fit)

#Considering strBlock as a predictor
fit2<- lm(Average ~ Hosp + strBlockint, data = df2)


#strBlock * Hosp interaction term
fit3<- lm(Average ~ Hosp * strBlockint, data = df2)


df2 |>
  group_by(strBlock, Hosp) |>
  summarize(hosp_entrustment_score = mean(Average))|>
ggplot(aes(x = strBlock, y = hosp_entrustment_score, color = Hosp)) + 
  geom_point() +
  geom_smooth(method = "lm",se = FALSE, aes(group = Hosp)) +
  scale_y_continuous(limits = c(1,5) ) +
  scale_color_manual(values = c("red", "blue")) +
  labs(title = "Comparison of Entrustment Scores Between Student Employees 
       & Student Non-Employees", x = "Rotation Block",
       y = "Entrustment Score", color = "Mean by Rotation") + 
  theme_minimal()
```
```{r}
#t-test to compare means between the two groups 
t.test(Average ~ Hosp, data = df2, var.equal = FALSE )

#highly significant!
```

robbie-work



## Comparison of Entrustment Scores Between Students Who Compeleted Introductory Rotation Courses vs Those Who Didn't
```{r}
#simple linear regression
df3<- df |>
  select(strBlock, Early, Average) |>
  mutate(Early = factor(Early, labels = c("No Introductory Course", "Introductory Rotation Course")))

labels <- c("A1", "A2", "A3", "A4", "A5", "A6", "A7", "B1", "B2", "B3", "B4", 
            "B5", "B6", "C1", "C2", "C3", "C4", "C5", "C6", "D1","D2", 
            "D3", "D4", "D5", "D6")

# create a mapping of labels to integers
label_mapping <- setNames(1:length(labels), labels)

# convert strBlock column to integers
df3$strBlockint<- label_mapping[df3$strBlock]

#Without considering strBlock as a predictor
fit<- lm(Average ~ Early, data = df3)
summary(fit)

#Considering strBlock as a predictor
fit2<- lm(Average ~ Early + strBlockint, data = df3)


#strBlock * Hosp interaction term
fit3<- lm(Average ~ Early * strBlockint, data = df3)


df3 |>
  group_by(strBlock, Early) |>
  summarize(early_entrustment_score = mean(Average))|>
ggplot(aes(x = strBlock, y = early_entrustment_score, color = Early)) + 
  geom_point() +
  geom_smooth(method = "lm",se = FALSE, aes(group = Early)) +
  scale_y_continuous(limits = c(1,5) ) +
  scale_color_manual(values = c("red", "blue")) + 
    labs(title = "Comparison of Entrustment Scores Between Students Who
       Completed Introductory Rotations vs Those Who Did Not", x = "Rotation Block",
       y = "Entrustment Score", color = "Mean by Rotation") +
  theme_minimal()
```
```{r}
#t-test to compare means between the two groups 
t.test(Average ~ Early, data = df3, var.equal = FALSE )

#no significant difference
```
main
