---
title: "AdditionalStudents_EDA"
author: "Melissa Eckert"
date: "2025-03-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(openxlsx)
```

Load Data
```{r}
df<- read.xlsx("../Data/OffCycleStudents.xlsx")
glimpse(df)
```

Check for missing values 
```{r}
df_missing <- df %>% filter(is.na(medical) | is.na(clinical_reasoning) | is.na(professionalism) | is.na(collaboration))
df_missing

#lets remove strBlock D2 for ID 243 because most values are NA
df<- df |> filter(!(ID == 243 & strBlock == "D2"))



```
When do students start their rotations?
```{r}
df|>
  filter(Number.Rotation == 1)|>
  ggplot(aes(x = strBlock)) +
  geom_bar(fill = "darkgreen") +
  labs(title = "Distribution of First Rotation Blocks",
       x = "First Rotation Block", y = "Number of Students") +
  scale_y_continuous(limits = c(0,15))+
  theme_minimal()
```
Most students in the data set still start at A1. So what makes them abnormal? 

```{r}
repeated_rotations <- df |>
  group_by(ID, strBlock) |>
  filter(n() > 1) |>
  distinct(ID, strBlock)
repeated_rotations
```

```{r}
#Look at average scores for all students by rotation order 
  df |>
  group_by(Number.Rotation, grad_year) |>
  summarise(average = mean(average)) |>
  ggplot(aes(x = Number.Rotation, y = average, color = grad_year)) +
  geom_line( size = 1) +
  geom_point() +
  labs(title = "Student Performance Over Rotation Order by Year",
       x = "Rotation Order",
       y = "Average Score") +
  facet_wrap(~grad_year, nrow = 2, ncol = 1)+
  theme_minimal()+
  theme(legend.position = "none")
```
Student Scores by Rotation Block
```{r}
 df |>
  group_by(strBlock, grad_year) |>
  summarise(average = mean(average)) |>
  ggplot(aes(x = strBlock, y = average, fill = grad_year)) +
  geom_bar(stat = "identity") +
  labs(title = "Student Performance by Rotation Block",
       x = "Rotation Block",
       y = "Average Score") +
  facet_wrap(~grad_year, nrow = 2, ncol = 1)+
  theme_minimal()+
  theme(legend.position = "none")
```

Do students who take their 9th rotation in block A1 take a major hit in assessment because new students start?


```{r}
#look at students who start at rotation block C5 (A1 = 8 or 9)
 C5_students <- df |>
  filter(Number.Rotation == 1 & strBlock == "C5") |>
  pull(ID) 

df_C5<- df |>
  filter(ID %in% C5_students)
df_C5|>
  group_by(Number.Rotation)|>
  summarize(average = mean(average))|>
  ggplot(aes(x = Number.Rotation, y = average)) +
  geom_point()+
  geom_line() +
  labs(title = "Performance Over Time for Students Starting at C5", x = "Rotation Number", 
       y = "Average Score") +
  theme_minimal()

#look at their performance by rotation block 
df_C5 |>
  group_by(strBlock) |>
  summarise(average = mean(average)) |>
  ggplot(aes(x = strBlock, y = average)) +
  geom_bar(stat = "identity", fill = "darkgreen") +
  labs(title = "Performance by Rotation Block for Students Starting at C5", 
       x = "Rotation Block", y = "Average Score")
  
```
There is a large dip at around rotation number 9 and 10, which is either A1, A2, A3, or A4.

**Questions**
1. How do we determine which students did not complete years 1-3 at Cornell?
  ID's for 2024 data go from 1-117
  ID's for 2025 data go from 118-234
  ID's for this data for from 235-263
  