---
title: "Multinomial Logit Models"
author: "Robert Feldstein"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(readxl)
library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)
library(nnet)
library(MASS)
library(ordinal)
```

```{r}
# We need both the 2025 data, and the procedure data

df25 <- read_xlsx("../Data/2025Data.xlsx", sheet = 1)
dfproc <- read_xlsx("../Data/FinalProcedures.xlsx", sheet = 1)

# We need to aggregate the procedure data to the student level
# For now, we will just look at Ovarioectomy/Ovariohysterectomy and Castration

neuters <- dfproc %>%
  filter(skill %in% c("Ovarioectomy/Ovariohysterectomy", "Castration")) %>%
  group_by(ID, skill) %>%
  summarise(total = sum(total_number_performed, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = skill, values_from = total, values_fill = 0, 
              names_prefix = "Total_")

# Rename neuters column names
colnames(neuters) <- c("ID", "Total_Ov", "Total_Castration")


# Add time variable

labels <- c("A1", "A2", "A3", "A4", "A5", "A6", "A7", "B1", "B2", "B3", "B4", 
            "B5", "B6", "C1", "C2", "C3", "C4", "C5", "C6", "D1","D2", 
            "D3", "D4", "D5", "D6")

# Create a mapping of labels to integers
label_mapping <- setNames(1:length(labels), labels)

# Convert strBlock column to integers
df25$time<- label_mapping[df25$strBlock]

# This maps missing catalog numbers to the last service for each student

last_rotation <- df25 %>%
  group_by(UniqueID) %>%
  summarise(last_service = last(service), .groups = "drop")
dfproc_course <- dfproc %>%
  filter(skill %in% c("Ovarioectomy/Ovariohysterectomy", "Castration")) %>%
  # join last rotation info based on student ID (which is called ID in dfproc)
  left_join(last_rotation, by = c("ID" = "UniqueID")) %>%
  # if catalog_number is missing, use the last_service from df25
  mutate(catalog_number_fixed = if_else(is.na(catalog_number), last_service, catalog_number)) %>%
  # select the updated column name
  dplyr::select(ID, catalog_number = catalog_number_fixed, skill, total_number_performed)
neuters <- dfproc_course %>%
  group_by(ID, catalog_number, skill) %>%
  summarise(total = sum(total_number_performed, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = skill, values_from = total, values_fill = 0, 
              names_prefix = "Total_")
colnames(neuters) <- c("ID", "catalog_number", "Ov", "Castration")

df25 <- df25 %>%
  left_join(neuters, by = c("UniqueID" = "ID", "service" = "catalog_number")) %>%
  mutate(
    Castration = replace_na(Castration, 0),
    Ov = replace_na(Ov, 0)
  )


df25$CastrationPerformed <- ave(df25$Castration, df25$UniqueID, FUN = cumsum)
df25$OvPerformed <- ave(df25$Ov, df25$UniqueID, FUN = cumsum)


skill_cat <- dfproc %>%
  left_join(last_rotation, by = c("ID" = "UniqueID")) %>%
  mutate(catalog_number_fixed = if_else(is.na(catalog_number), last_service, catalog_number)) %>%
  dplyr::select(ID, catalog_number = catalog_number_fixed, skill_category, total_number_performed) %>%
  group_by(ID, catalog_number, skill_category) %>%
  summarise(total = sum(total_number_performed, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = skill_category, values_from = total, values_fill = 0)

df25 <- df25 %>% 
  left_join(skill_cat, by = c("UniqueID" = "ID", "service" = "catalog_number")) %>%
  mutate(
    `Surgical Skills` = replace_na(`Surgical Skills`, 0),
    `Procedural Skills` = replace_na(`Procedural Skills`, 0),
    `Dental Skills` = replace_na(`Dental Skills`, 0),
    `Anesthesia Skills` = replace_na(`Anesthesia Skills`, 0)
  )

df25$Surgical_Per <- ave(df25$`Surgical Skills`, df25$UniqueID, FUN = cumsum)
df25$Procedural_Per <- ave(df25$`Procedural Skills`, df25$UniqueID, FUN = cumsum)
df25$Dental_Per <- ave(df25$`Dental Skills`, df25$UniqueID, FUN = cumsum)
df25$Anesthesia_Per <- ave(df25$`Anesthesia Skills`, df25$UniqueID, FUN = cumsum)
```


```{r}

# Convert to factors

df25$medical <- as.factor(df25$medical)
df25$clinical_reasoning <- as.factor(df25$clinical_reasoning)
df25$professionalism <- as.factor(df25$professionalism)
df25$collaboration <- as.factor(df25$collaboration)
df25$patient_care <- as.factor(df25$patient_care)
df25$procedural <- as.factor(df25$procedural)

```

```{r}
# Basic Multinomial Model
mnl_model <- multinom(medical~OvPerformed+ CastrationPerformed + time + PIPOverall,data = df25)
summary(mnl_model)
```


```{r}
# Ordered logit assumes that the categories are ordered (e.g. 1 < 2 < 3 < 4 < 5)
# Fit the model (logistic has lowest AIC)
ordered_medical <- polr(medical ~ OvPerformed + CastrationPerformed + time + PIPOverall, data = df25, Hess = TRUE,
                        method = "logistic")
summary(ordered_medical)
```

Interpretation: 

Note: P-values are missing, so we look at the t value. If the magnitude of the t-value is greater than 1.96, this would equate to a p-value less than 0.05, indicating significance.

Coefficients: Each of the coefficients represents the log odds of moving up one level in the ordered factor. For example, a positive coefficient for OvPerformed would indicate that as the number of Ovarioectomies performed increases, the odds of moving up a level in the medical rating also increase. We would take the exponential of the coefficient to get the precise odds ratio.

Intercepts: These represent the cutoff points between the different levels of the ordered factor. For example, the first intercept is the cutoff between the first and second levels, the second intercept is the cutoff between the second and third levels, and so on. These cutoffs are floating point numbers because they represent the cut off for the latent variable that is being estimated, not the visible ordered factor.The magnitude of the intercept can be interpreted as the "size" of the gap between the levels. The more negative an intercept, the less likely a student is to be in that level gap. 


```{r}
# Lowest AIC is loglog
ordered_clinical <- polr(clinical_reasoning ~ OvPerformed + CastrationPerformed + time + PIPOverall, data = df25, Hess = TRUE, method = "logistic")
summary(ordered_clinical)
```


```{r}
ordered_professionalism <- polr(professionalism ~ OvPerformed + CastrationPerformed + time + PIPOverall, data = df25, Hess = TRUE, method = "logistic")
summary(ordered_professionalism)
```


```{r}
ordered_collaboration <- polr(collaboration ~ OvPerformed + CastrationPerformed + time + PIPOverall, data = df25, Hess = TRUE, method = "logistic")
summary(ordered_collaboration)
```


```{r}
ordered_patient <- polr(patient_care ~ OvPerformed + CastrationPerformed + time + PIPOverall, data = df25, Hess = TRUE, method = "logistic")
summary(ordered_patient)
```

```{r}
ordered_procedural <- polr(procedural ~ OvPerformed + CastrationPerformed + time + PIPOverall, data = df25, Hess = TRUE, method = "logistic")
summary(ordered_procedural)
```

## 2024 Multinomial Logit Models

```{r}
df24 <- read_xlsx("../Data/FinalDataSet.xlsx", sheet = 1)

# Convert strBlock column to integers
df24$time<- label_mapping[df24$strBlock]

df24$medical <- as.factor(df24$medical)
ordered_medical_24 <- polr(medical ~ Hosp + Early + PIPYN + time, data = df24)
```

```{r}


m1 <- clm(medical ~ Hosp + PIPActive + time, data = df24)
m2 <- clm(medical ~ Hosp + PIPActive + time, nominal = ~Hosp  + PIPActive + time, data = df24)
anova(m1, m2)

```

Unforunately, one of the key assumptions of POLR is violated. So we can use a non-ordinal multinomial logistic regression model instead, which does not assume proportional effects of variables.


```{r}
df24$medical_factor <- factor(df24$medical)
df24$medical_factor <- relevel(df24$medical_factor, ref = "3")

model <- multinom(medical_factor ~ Hosp + Early + PIPActive + time, data = df24)
summary(model)
```





Here is a more advanced function that can automatically calculate the p-values for us, which is a nice help.

```{r}
# We set the reference level to 3, since that helps us compare competency vs. non-competency
multinom_logit <- function(var, df, predictors = NULL, ref_level = "3") {
  # Start by making the model formula
  if (is.null(predictors)) {
    stop("Please provide a vector of predictor variable names")
  }
  # Create a copy of the dataframe so we do not mess with the existing data
  df_copy <- df
  # Set the reference level
  var_factor_name <- paste0(var, "_factor")
  df_copy[[var_factor_name]] <- factor(df_copy[[var]])
  # Check if reference level exists in the data (should be fine, except for values of 1)
  if (!(ref_level %in% levels(df_copy[[var_factor_name]]))) {
    stop(paste("Reference level", ref_level, "not found in the data"))
  }
  df_copy[[var_factor_name]] <- relevel(df_copy[[var_factor_name]], ref = ref_level)
  # Make the formula string (thank you Professor Entner)
  formula_str <- paste(var_factor_name, "~", paste(predictors, collapse = " + "))
  formula_obj <- as.formula(formula_str)
  # Fit the model
  # set trace to false, so that way we are not flooded with the convergence information
  model <- multinom(formula_obj, data = df_copy, trace = FALSE)
  # Calculate z-statistics using typical value - 0/(sqrt(var))
  z_stats <- coef(model) / summary(model)$standard.errors
  # Calculate p-values with pnorm function (don't forget to double it)
  p_values <- (1 - pnorm(abs(z_stats))) * 2
  # Format p-values for added interpretability
  get_stars <- function(p) {
    stars <- ""
    if (p < 0.001) stars <- "***"
    else if (p < 0.01) stars <- "**"
    else if (p < 0.05) stars <- "*"
    else if (p < 0.1) stars <- "."
    return(stars)
  }
  p_values_formatted <- matrix(
    paste0(format(round(p_values, 4), nsmall = 4), 
           " ", 
           apply(p_values, c(1, 2), get_stars)),
    nrow = nrow(p_values)
  )
  rownames(p_values_formatted) <- rownames(p_values)
  colnames(p_values_formatted) <- colnames(p_values)
  
  # Calculate odds ratios
  odds_ratios <- exp(coef(model))
  
  # Create custom summary table
  coefs <- coef(model)
  std_errors <- summary(model)$standard.errors
  
  result_summary <- list()
  for (i in 1:nrow(coefs)) {
    category <- rownames(coefs)[i]
    result_summary[[category]] <- data.frame(
      Coefficient = coefs[i,],
      StdError = std_errors[i,],
      OddsRatio = odds_ratios[i,],
      p_value = p_values[i,],
      Significance = apply(matrix(p_values[i,], nrow = 1), 2, get_stars)
    )
  }
  
  # Run an overall likelihood ratio test for each predictor
  lr_results <- data.frame(
    Predictor = character(),
    LR_Stat = numeric(),
    df = numeric(),
    p_value = numeric(),
    stringsAsFactors = FALSE
  )
  
  # Full model
  full_model <- model
  
  # Test each predictor
  for (pred in predictors) {
    # Create reduced formula without this predictor
    reduced_preds <- predictors[predictors != pred]
    reduced_formula <- as.formula(paste(var_factor_name, "~", 
                                       ifelse(length(reduced_preds) > 0,
                                              paste(reduced_preds, collapse = " + "),
                                              "1")))
    
    # Fit reduced model
    reduced_model <- tryCatch({
      multinom(reduced_formula, data = df_copy, trace = FALSE)
    }, error = function(e) {
      message("Error fitting reduced model without ", pred, ": ", e$message)
      return(NULL)
    })
    
    if (!is.null(reduced_model)) {
      # Calculate likelihood ratio test
      lr_stat <- 2 * (logLik(full_model) - logLik(reduced_model))
      degrees_freedom <- length(coef(full_model)) - length(coef(reduced_model))
      p_val <- 1 - pchisq(lr_stat, degrees_freedom)
      
      # Add to results
      lr_results <- rbind(lr_results, data.frame(
        Predictor = pred,
        LR_Stat = as.numeric(lr_stat),
        df = degrees_freedom,
        p_value = as.numeric(p_val),
        stringsAsFactors = FALSE
      ))
    }
  }
  
  # Add significance markers
  if (nrow(lr_results) > 0) {
    lr_results$Significance <- sapply(lr_results$p_value, get_stars)
  }
  
  # Return all results in a list
  return(list(
    model = model,
    summary = summary(model),
    p_values = p_values_formatted,
    odds_ratios = odds_ratios,
    coefficient_summary = result_summary,
    likelihood_ratio_tests = lr_results,
    formula = formula_obj,
    reference_level = ref_level
  ))
}

# Pretty print function to display all results nicely
print_multinom_results <- function(results) {
  cat("=== MULTINOMIAL LOGISTIC REGRESSION RESULTS ===\n")
  cat("\nModel Formula:", deparse(results$formula), "\n")
  cat("Reference Level:", results$reference_level, "\n\n")
  
  cat("=== COEFFICIENTS AND P-VALUES ===\n")
  for (cat_name in names(results$coefficient_summary)) {
    cat("\nCategory", cat_name, "vs Reference", results$reference_level, ":\n")
    print(results$coefficient_summary[[cat_name]])
  }
  
  cat("\n=== LIKELIHOOD RATIO TESTS FOR PREDICTORS ===\n")
  if (nrow(results$likelihood_ratio_tests) > 0) {
    print(results$likelihood_ratio_tests)
  } else {
    cat("No likelihood ratio tests could be computed.\n")
  }
  
  cat("\n=== MODEL FIT ===\n")
  cat("Residual Deviance:", results$model$deviance, "\n")
  cat("AIC:", results$model$AIC, "\n")
  cat("Number of observations:", results$model$edf, "\n")
}

```

```{r}
medical_24 <- multinom_logit(
  var = "medical", 
  df = df24, 
  predictors = c("Hosp", "Early", "PIPYN", "time")
)
print_multinom_results(medical_24)
```

```{r}

# Remove category of 1 from clinical reasoning
library(dplyr)
# filter 
df24 = df24 %>% filter(clinical_reasoning!=1) 

clinical_24 <- multinom_logit(
  var = "clinical_reasoning", 
  df = df24, 
  predictors = c("Hosp", "Early", "PIPActive", "time", "procedural")
)
print_multinom_results(clinical_24)
```

```{r}

colab_24 <- multinom_logit(
  var = "collaboration", 
  df = df24, 
  predictors = c("Hosp", "Early", "PIPActive", "time")
)
print_multinom_results(colab_24)

```

```{r}
prof_24 <- multinom_logit(
  var = "professionalism", 
  df = df24, 
  predictors = c("Hosp", "Early", "PIPActive", "time")
)
print_multinom_results(prof_24)

```

```{r}
clinical_25 <- multinom_logit(
  var = "clinical_reasoning",
  df = df25,
  predictors = c("OvPerformed", "CastrationPerformed", "time", "PIPOverall")
)
print_multinom_results(clinical_25)

```

In the future, we might be able to make an R-Shiny visualization out of this.

```{r}
# Run a quick PCA on the df24
# Let medical score be the outcome variable 

df24$medical <- as.numeric(df24$medical)
df24$clinical_reasoning <- as.numeric(df24$clinical_reasoning)
df24$professionalism <- as.numeric(df24$professionalism)
df24$collaboration <- as.numeric(df24$collaboration)

pca <- prcomp(~medical + clinical_reasoning + 
            professionalism + collaboration + patient_care + procedural, data = df24, scale = T)
summary(pca)

pca$rotation
biplot(pca, scale = 0)
```
Look at weights to understand the pairing of variables; we see that medical and clinical_reasoning are paired together, while professionalism and collaboration.
