---
title: "Detrended Linear Regression Models"
author: "Robert Feldstein"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(nnet)
```


### 2024 Data

```{r}

df24 <- read_xlsx("../Data/FinalDataSet.xlsx", sheet = 1)
labels <- c("A1", "A2", "A3", "A4", "A5", "A6", "A7", "B1", "B2", "B3", "B4", 
            "B5", "B6", "C1", "C2", "C3", "C4", "C5", "C6", "D1","D2", 
            "D3", "D4", "D5", "D6")

# Create a mapping of labels to integers
label_mapping <- setNames(1:length(labels), labels)

# Convert strBlock column to integers
df24$time<- label_mapping[df24$strBlock]

```

### 2024 Medical Detrended

```{r}

df24_complete <- df24[complete.cases(df24$medical, df24$time), ]

# Fit the linear model on the complete cases
lm24.med <- lm(medical ~ time, data = df24_complete)

# Extract residuals
df24_complete$medresid <- residuals(lm24.med)

# Merge residuals back into the original dataframe
df24$medresid <- NA  # Initialize with NA
df24$medresid[complete.cases(df24$medical, df24$time)] <- df24_complete$medresid

# Now fit a linear regression model to these residuals.

lm24.res <- lm(medresid ~ Hosp + PIPYN + Early, data = df24)
summary(lm24.res)

```

## Ordinal Logistic Regression

```{r}
library(MASS)

# Convert medical to an ordered factor
df24$medical <- factor(df24$medical, ordered = TRUE)

# Fit an ordinal logistic regression model
ord_logit <- polr(medical ~ Hosp + Early + PIPYN + time, data = df24, method = "logistic")
summary(ord_logit)


```

Interpretation:

This model is actually easier to interpret than the previous model we utilized. The outputted coefficients are in terms of log-odds. They represent the change in log-odds for each one unit increase in the variable. For PIPYN, a value of -0.89 in log-odds means that students in a PIP are 41% as likely to score higher than students not in a PIP. For time, a log odds of around 0.1 means that for every rotation block, a student is about 1.1 times as likely to score higher than they did in the previous rotation block. This compounds overtime, after 10 rotations a student is 2.6 times as likely to score higher than they did at the start.

The intercepts represent the log-odds of being at or below the cutoff point, before the variables are considered. E.g. what are the chances of being at a 3 or lower regardless of a student's individual attributes. The higher the value, the lower the probability of being a specific number.

## 2025 Data


```{r}

df25 <- read_xlsx("../Data/2025Data.xlsx", sheet = 1)

# Convert strBlock column to integers
df25$time<- label_mapping[df25$strBlock]

```


```{r}
# Try fitting a random effect on studentID 

library(lme4)
model <- lmer(medical ~ time + Hosp + Early + PIPYN + (1 | UniqueID), data = df24)
summary(model)

```

```{r}

# Cumulative Linked Mixed Models 
library(ordinal)
df24$medical_f <- as.factor(df24$medical)
clmodel <- clmm(medical_f ~ time + Hosp + Early + PIPYN + (1 | UniqueID), data = df24, Hess = T)
summary(clmodel)

```


```{r}
# Try a simplified DID model here 
# Average of rotations 1-7, average of rotations 7-end

df24$post <- ifelse(df24$time >12, 1, 0)
df24$treat <- ifelse(df24$PIPYN == 1, 1, 0)


# Create interaction term
df24$did <- df24$post * df24$treat

# Run DID regression
model <- lm(clinical_reasoning ~ post + treat + did, data = df24)
summary(model)


```

```{r}
library(fixest)
feols(clinical_reasoning ~ post + treat + did, cluster = ~UniqueID, data = df24)
```

```{r}
# Check the assumptions for POLR for df24
# medical ~ hosp + PIPYN + early + time 

sf <- function(y) {
  c('Y>=1'= qlogis(mean(y>=1)),
    'Y>=2' = qlogis(mean(y >=2)),
    'Y>=3' = qlogis(mean(y>=3)),
    'Y>=4' = qlogis(mean(y>=4)),
    'Y>=5' = qlogis(mean(y>=5))
    )
}

(s <- with(df24, summary(as.numeric(medical) ~ Hosp + PIPYN + Early + time, fun = sf)))

glm(I(as.numeric(medical) >= 4) ~ Hosp, family="binomial", data = df24)
glm(I(as.numeric(medical) >= 3) ~ Hosp, family="binomial", data = df24)

```

