---
title: "Procedure Data"
author: "Robert Feldstein"
date: "2025-02-26"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(openxlsx)
library(dplyr)

df <- read.xlsx("../Data/Procedures.xlsx", sheet = 1)
df_new <- read.xlsx("../Data/FinalProcedures.xlsx", sheet = 1)
# Rename df_new catalog column to be called service

colnames(df_new)[colnames(df_new) == "catalog_number"] <- "service"
colnames(df_new)[colnames(df_new) == "skill"] <- "Skill"
colnames(df_new)[colnames(df_new) == "ID"] <- "Unique.ID"

```


```{r}
# Potential Dataset issues 
# Look at the NA rows

missing<- df_new %>% filter(is.na(service) |service < 1000)

# There are 250 rows where there is no service marked 

# Check about total_number_performed being missing as well

df_new %>% filter(is.na(total_number_performed))

# 211 rows where it is just empty in the count
# It is allowed to just replace these rows with zeros

```

## Data Cleaning

```{r}
# First clean up service values that are less than 1000 by inserting them into the total_number_performed category
df_new$service <- as.numeric(df_new$service)
df_new <- df_new %>% mutate(total_number_performed = ifelse(service < 1000, service, total_number_performed))
# Now set any service less than 1000 to "Missing"

df_new <- df_new %>% mutate(service = ifelse(service < 1000 | is.na(service), "Missing", service))
df_new$service <- as.factor(df_new$service)

# Set any NA values in total_number_performed to zero

df_new$total_number_performed[is.na(df_new$total_number_performed)] <- 0

```



```{r}
# Castration and Ovarioectomy are the most important 

# Group by skill and get the count 

df_new %>% group_by(Skill,grad_year) %>% summarise(Count = sum(total_number_performed)) %>% arrange(desc(Count))

```

Castration and Ovarioectomy are the most popular procedures. They will be important to track.



```{r}
# Group by skill and UniqueID

df_new %>% group_by(Unique.ID, Skill) %>% summarise(Count = sum(total_number_performed))

```

This is an individual list of procedures performed by each student. 

## Investigating Differences between 2024 and 2025 

```{r}


# How about sum of count column

df_new %>% group_by(grad_year) %>% summarise(Count = sum(total_number_performed, na.rm = TRUE))

```

As known by our client, students did not log nearly as many procedures in 2024 as they did in 2025. This is largely because the procedural tracker was not introduced until January. There were essentally 5000 more procedures logged in 2025 than in 2024. 



```{r}
# Which services have the most surgeries?

df_new %>% group_by(service,grad_year) %>% summarise(Count = sum(total_number_performed, na.rm = TRUE)) %>% arrange(desc(Count))

```
The class with the most surgeries is 5606.


```{r}
# What are the most popular categories of surgeries?

df_new %>% group_by(skill_category,grad_year) %>% summarise(Total = sum(total_number_performed, na.rm = TRUE), Unique = n()) %>% arrange(desc(Total))


```

```{r}
# Come up with some plots

library(ggplot2)

# Procedures by category sum, but break the colors into 2024 and 2025

# ggplot(df, aes(x = Category, fill = Class)) + geom_bar() + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ggtitle("Total Counts of Procedures for 2024 and 2025 (C6)")

# Do the same but sum up the count column 
df_new$grad_year <- as.factor(df_new$grad_year)

ggplot(df_new, aes(x = skill_category, y = total_number_performed, fill = grad_year)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ggtitle("Total Counts of Procedures for 2024 and 2025 (C6)")



```

```{r}
# Procedures by service, but break the colors into 2024 and 2025

# ggplot(df, aes(x = service, fill = Class)) + geom_bar() + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ggtitle("Procedures by Class for 2024 and 2025 (C6)")

# Do this but sum the count

df_new$service <- as.factor(df_new$service)

ggplot(df_new, aes(x = service, y = total_number_performed, fill = grad_year)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ggtitle("Procedures by Class for 2024 and 2025 (C6)")

```


```{r}
# Plot a distribution for each category

ggplot(df_new, aes(x = skill_category, y = total_number_performed)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
# Do a distribution by category for 2024 and 2025

ggplot(df_new, aes(x = skill_category, y = total_number_performed, fill = grad_year)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 45, hjust = 1))  + ggtitle("Boxplots of Procedures by Category for 2024 and 2025 (C6)")

```


```{r}
# Plot a barchart for the average number of procedures by category

df_new %>% group_by(skill_category) %>% summarise(Avg = mean(total_number_performed, na.rm = TRUE)) %>% ggplot(aes(x = skill_category, y = Avg)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ggtitle("Average Procedures by Category (C6)")

```



## Analysis of Neuters 

```{r}
# There is a value to just looking at the castrations and ovariectomies between the populations

neuters <- df_new %>% filter(Skill %in% c("Castration", "Ovarioectomy/Ovariohysterectomy"))

# Group by grad_year and skill

neuters %>% group_by(grad_year, Skill) %>% summarise(Count = sum(total_number_performed, na.rm = TRUE))

# Make a histogram of the total number of neuters

ggplot(neuters, aes(x = total_number_performed, fill = grad_year)) + geom_histogram(binwidth = 1) + ggtitle("Total Number of Neuters by Year (C6)")

# Make a boxplot that breaks into castration and ovarioectomy

ggplot(neuters, aes(x = Skill, y = total_number_performed, fill = grad_year)) + geom_boxplot() + ggtitle("Boxplot of Neuters by Year (C6)")


```



## Cluster the Population into those who report a lot and those who do not

```{r}

# Subset neuters into 95% and 5% populations

neuters_subset <- neuters %>% filter(total_number_performed < quantile(neuters$total_number_performed, 0.9, na.rm = TRUE))

# Make a histogram of the total number of neuters

ggplot(neuters_subset, aes(x = total_number_performed, fill = grad_year)) + geom_histogram(binwidth = 1) + ggtitle("Total Number of Neuters by Year (C6)")

# Make a boxplot that breaks into castration and ovarioectomy

ggplot(neuters_subset, aes(x = Skill, y = total_number_performed, fill = grad_year)) + geom_boxplot() + ggtitle("Boxplot of Neuters by Year (C6)")



```



